---
title: "data.world R-SDK Story"
output: html_notebook
---

Intro to data.world r-sdk

* Preconditions:
    + a data.world account
    + R binary 

1. Connect to data.world 
Insert your token into a `.data.world` file in your home directory e.g 
``` {bash from your terminal}
echo 'token=YOUR_API_TOKEN' > ~/.data.world
```
(You can find your API token at https://data.world/settings/advanced)
```{r setup connection}
# load library
library("data.world")
conn <- data.world()
userId <- YOUR_DATAWORLD_USERID
```

2. View the contents of a public dataset
```{r getDataset, echo=TRUE}
cancer_dataset <- data.world::getDataset(connection = conn, dataset = "nrippner/cancer-trials")
cancer_dataset$files
```
or query a dataset
```{r}
query(connection = conn, type = "sql", dataset = "nrippner/cancer-trials", query = "SELECT * FROM Tables")
```

3. Create a new dataset right from your R environment to publish / share your  finding
```{r createDataset, echo=TRUE}
sample_dataset_title = "sampledataset"
sample_dataset_id = sprintf("%s/%s", userId, sample_dataset_title)
createDataset(
  connection = conn,
  createDatasetRequest = data.world::DatasetCreateRequest(
      title = sample_dataset_title,
      visibility = "PRIVATE",
      description = "your first dataset created by R-SDK" ,
      tags = c("rsdk", "sdk") ,
      licenseString = "Public Domain"),
  ownerId = userId
  )
```

4. Upload any files to a dataset to which the user is a contributor
```{r uploadFile, echo=TRUE}
uploadFile(connection = conn, path = "README.md", fileName = "README.md", sample_dataset_id)
```

5. Query any public dataset and store that in a data frame
```{r queryDataset, echo=TRUE}
cancer_trials_tables <- query(connection = conn, type = "sql", dataset = "nrippner/cancer-trials", query = "SELECT * FROM Tables")
cancer_trials_tables
```

6. Upload that data frame to another dataset
```{r uploadDataFrame, echo = TRUE}
uploadDataFrame(connection = conn, fileName="cancer_trials_table.csv", dataFrame = cancer_trials_tables, dataset = sample_dataset_id)
```

7. Suppose that user wants to do a more involved calculation with a data file (maybe because the dwsql/sparql query maybe too complicated) . The user can download the file locally
```{r downloadFile, echo = TRUE}
downloadFile(connection = conn, dataset = "nrippner/cancer-trials", fileName = "cen_income.csv", output = "tmp/cen_income.csv")
```
or the user can download that file directly into a data frame and perform additional operations
```{r downloadFileAsDataFrame}
cancer_cen_income <- downloadFileAsDataFrame(connection = conn, dataset = "nrippner/cancer-trials", fileName = "cen_income.csv")
coffee_county_cen_income <- subset(cancer_cen_income, Name == 'Coffee County', select=c(Name, Postal.Code, Median.Household.Income))
coffee_county_cen_income
```
and upload to the dataframe
```{r uploadDataframe, echo=TRUE}
uploadDataFrame(connection = conn, fileName="coffee_county_cen_income.csv", dataFrame = coffee_county_cen_income, dataset = sample_dataset_id)
```

```{r verify, echo=TRUE}
downloadFileAsDataFrame(connection = conn, dataset = sample_dataset_id, fileName = "coffee_county_cen_income.csv")
```

8. User can collect, import(link) similar datas from the web to this dataset (discovered via https://data.world/opendata/data.gov for example)
```{r addFileBySource, echo=TRUE}
addFileBySource(connection = conn, dataset = sample_dataset_id, name = "alabama_gravesites.csv", url = "https://inventory.data.gov/dataset/1cf2471d-cfc6-4004-9d98-1e8d3e9cb5ae/resource/868c08d1-27c9-41f9-b486-d7c24dfd4aa9/download/nglalabama.csv")

```

9. Clean up / delete file(s)
```{r deleteFileAndSyncSources, echo=TRUE}
deleteFileAndSyncSources(connection = conn, dataset = sample_dataset_id, name = "coffee_county_cen_income.csv")
```

10. Publish your dataset
```{r patchDataset, echo=TRUE}
publish_dataset_request <- data.world::DatasetPatchRequest(visibility = "OPEN")
data.world::patchDataset(connection = conn, datasetPatchRequest = publish_dataset_request, datasetid = sample_dataset_id)
```
