<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{data.world datapackage guide}
-->

```{r, echo = FALSE}
library(data.world)
```

# datapackage
The goal of this document is to give a quick introduction to the datapackage construct and how it should be used. 

## Setting up the connection
Once installed, you can instantiate a connection by providing your
data.world API token to the constructor.  You can find your API
token at https://data.world/settings/advanced

you can insert your token into a `.data.world` file in your home
directory, and the constructor will read it from there:
```{bash from your terminal, eval = FALSE}
echo 'token=YOUR_API_TOKEN' > ~/.data.world
```
then
```{r}
conn <- data.world()
```

## Datapackage
datapackage is a standardized way to package data for simple sharing between tools and people.
Learn more about the Data Package specification at frictionlessdata.io.

To download a datapackage 
```{r downloadDatapackage, eval = FALSE}
introDataPackage <- data.world::downloadDatapackage(connection = data.world() , dataset = "jonloyens/an-intro-to-dataworld-dataset")
```

To list tables in a datapackage
```{r listDatapackageTables, eval = FALSE}
introDataPackagetables <- data.world::listTables(introDataPackage)
introDataPackagetables
```

To load a datapackage table as a dataframe
```{r loadDatapackageTable, eval = FALSE}
changeLogTable <- data.world::loadTable(introDataPackage, 'changelog')
```

To view a table schema
```{r viewDatapackageTableSchema, eval = FALSE}
changeLogTableSchema <- data.world::loadSchema(introDataPackage, 'changelog')
changeLogTableSchema
```
