---
title: "data.world query"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{query}
  %\VignetteEncoding{UTF-8}
---
```{r, echo = FALSE}
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
CONFIGURED <- !is.null(getOption("dw.auth_token"))
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "##",
  purl = NOT_CRAN && CONFIGURED,
  eval = NOT_CRAN && CONFIGURED
)
```

# Querying

Datasets on data.world can be queried using SQL and SPARQL. A specific 
dataset is the starting point for a query, but its not the end. The data 
in that dataset can be joined with data from any other dataset on the platform.

If you are unfamiliar with or interested in learning more about SQL and SPARQL,
check out our tutorials:

* dwSQL: https://docs.data.world/tutorials/dwsql/
* SPARQL: https://docs.data.world/tutorials/sparql/


# SQL queries

## Inspecting the schema

When using SQL, it's a good idea to start by understanding the schema inferred
by data.world from the data that has been uploaded. That is 
done by querying the `Tables` table.

```{r tables}
intro_ds <- "https://data.world/jonloyens/an-intro-to-dataworld-dataset"
tables_df <- data.world::query(
  data.world::qry_sql("SELECT * FROM Tables"),
  dataset = intro_ds
)
tables_df
```

It is also possible to obtain information about columns. For that,
use the `TableColumns` table instead.

Try:
```{r columns, eval = FALSE}
columns_df <- data.world::query(
  data.world::qry_sql("SELECT * FROM TableColumns"),
  dataset = intro_ds
)
columns_df
```

## SELECT queries

Just like the queries above, any SQL query can be created using `qry_sql`.  
For example:
```{r SQL select}
assists_vs_height <- data.world::qry_sql(paste0(
  "SELECT t.Name, t.Height, s.AssistsPerGame ",
  "FROM DataDotWorldBBallTeam as t ",
  "JOIN DataDotWorldBBallStats as s ON t.Name = s.Name ",
  "ORDER BY s.AssistsPerGame DESC"))

data.world::query(assists_vs_height, dataset = intro_ds)
```

## Parameterized queries
SQL queries can be parameterized for maximum flexibility. Parameters are always
positional and can be provided via the `qry_sql` object.  
For example:
```{r SQL parameterized}
assists_greater_than <- data.world::qry_sql(paste0(
  "SELECT t.Name, t.Height, s.AssistsPerGame ",
  "FROM DataDotWorldBBallTeam as t ",
  "JOIN DataDotWorldBBallStats as s ON t.Name = s.Name ",
  "WHERE s.AssistsPerGame > ? ",
  "ORDER BY s.AssistsPerGame DESC"))

assists_greater_than$params <- c(10)
data.world::query(assists_greater_than, intro_ds)
```

# SPARQL

Running SPARQL queries is very similar to running SQL queries, except that
the `qry_sparql` constructor is used instead.

Additionally, instead of a tabular schema, with SPARQL one queries a data graph.

## SELECT queries

The previous SELECT query from the SQL example, translated to SPARQL, 
looks like this:
```{r SPARQL select}
assists_vs_height <- data.world::qry_sparql(paste(
  "BASE <http://data.world/jonloyens/an-intro-to-dataworld-dataset/> ",
  "PREFIX t: <DataDotWorldBBallTeam.csv/DataDotWorldBBallTeam#> ",
  "PREFIX s: <DataDotWorldBBallStats.csv/DataDotWorldBBallStats#> ",
  "SELECT ?name ?height ?assists WHERE { ",
  "  ?pt t:Name ?name . ",
  "  ?ps s:Name ?name . ", # the join column 
  "  ?pt t:Height ?height . ",
  "  ?ps s:AssistsPerGame ?assists . ",
  "} ",
  "ORDER BY DESC (?assists)", sep = "\n"
))

data.world::query(assists_vs_height, intro_ds)
```

## Parameterized SPARQL queries

Similarly, the previous parameterized query from the SQL example can
be translated to SPARQL as follows:
```{r SPARQL parameterized, eval = FALSE}
assists_greater_than <- data.world::qry_sparql(paste(
  "BASE <http://data.world/jonloyens/an-intro-to-dataworld-dataset/> ",
  "PREFIX t: <DataDotWorldBBallTeam.csv/DataDotWorldBBallTeam#> ",
  "PREFIX s: <DataDotWorldBBallStats.csv/DataDotWorldBBallStats#> ",
  "SELECT ?name ?height ?assists WHERE { ",
  "  ?pt t:Name ?name . ",
  "  ?ps s:Name ?name . ", # the join column 
  "  ?pt t:Height ?height . ",
  "  ?ps s:AssistsPerGame ?assists . ",
  "  FILTER(?assists > $v1) ",
  "} ",
  "ORDER BY DESC (?assists)", sep = "\n"
))
assists_greater_than$params <- c("$v1" = 10)
data.world::query(assists_greater_than, intro_ds)
```

Notice that in the SPARQL case, parameters are always named.

## Next

Queries are a powerful feature and are a great tool for pulling and wrangling data.  
For additional examples, check out our tutorials and published queries
[here](https://data.world/jonloyens/an-intro-to-dataworld-dataset/query) and
[here](https://data.world/jonloyens/intermediate-data-world/query).
