<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{data.world quickstart guide}
-->

```{r, echo = FALSE}
library(data.world)
```

# data.world quickstart guide
The goal of this document is to give a quick guide on how to use the r-sdk
to interact with data.world platform. 

## Setting up the connection
Once installed, you can instantiate a connection by providing your
data.world API token to the constructor.  You can find your API
token at https://data.world/settings/advanced

you can insert your token into a `.data.world` file in your home
directory, and the constructor will read it from there:
```{bash from your terminal, eval = FALSE}
echo 'token=YOUR_API_TOKEN' > ~/.data.world
```
then
```{r}
conn <- data.world()
```

## Create Dataset
To create a basic dataset with some metadata
```{r createDataset, eval = FALSE}
request <- data.world::DatasetCreateRequest(
      title = "sampledataset",
      visibility = "PRIVATE",
      description = "your first dataset created by R-SDK" ,
      tags = c("rsdk", "sdk") ,
      licenseString = "Public Domain",
      fileCreateRequests = list(data.world::FileCreateRequest("file1", "url1"))
      )

# Alternatively, you can also use data.world::addFile to add addtional files after the initial create construct
request <- data.world::addFile(request = request, name = "file4.csv",
      url = "https://data.world/file4.csv")

data.world::createDataset(
  connection = conn,
  createDatasetRequest = request,
  ownerId = "userid"
  )
```
*(Note that `title` and `visibility` are required and `visibility` has to be either `PRIVATE` or `OPEN`)*

## Replace Dataset
To replace an existing dataset
```{r replaceDataset, eval = FALSE}
request <- data.world::DatasetPutRequest(visibility = "OPEN", 
                                                   description = "new description", 
                                                   files = list(data.world::FileCreateOrUpdateRequest("file1", "url1")))

# Alternatively, you can also use data.world::addFile to add addtional files after the initial create construct
request <- data.world::addFile(request = request, 
                                         name = "file4.csv",
                                         url = "https://data.world/file4.csv")

data.world::replaceDataset(connection = conn, request, "userid/datasetid")
```

## Patch Dataset
To update an existing dataset 
```{r updateDataset, eval = FALSE}
request <- data.world::DatasetPatchRequest(visibility = "OPEN",
                                           description = "UPDATED DESCRIPTION !")

data.world::patchDataset(connection = conn,
                         datasetPatchRequest = request, 
                         datasetid = "userid/datasetid")
```

## Add Files by Source
Publish a single file on the web to be added to a data.world dataset
```{r addFileBySource, eval = FALSE}
data.world::addFileBySource(connection, 
                            dataset = "ownerid/datasetid",
                            name = "file.csv", 
                            url = "https://data.world/some_file.csv")
```
To fetch the latest files from source and update dataset after the initial uploading 
```{r sync, eval = FALSE}
data.world::sync(connection, dataset="ownerid/dataset")
```

## Delete File 
To delete a single file from a dataset
```{r deleteFileAndSyncSources, eval = FALSE}
data.world::deleteFileAndSyncSources(connection, 
                                     dataset = "ownerid/datasetid", 
                                     name = "file.csv")
```
To delete multiple files from a dataset
```{r deleteFilesAndSyncSources, eval = FALSE}
data.world::deleteFilesAndSyncSources(connection, 
                                      dataset = "ownerid/datasetid",
                                      names = list("file1.csv", "file2.csv"))
```

## Upload File
To upload one or more files a dataset
```{r uploadFiles, eval = FALSE}
data.world::uploadFiles(connection = conn, 
                        dataset = "ownerid/datasetid", 
                        paths = list ("file1.csv", "file2.csv"))
```
*(Note `paths` is an array of local file path relative to `getwd()`)*

## Upload Data Frame
To upload a dataframe to a data.world dataset as csv
```{r uploadDataFrame, eval = FALSE}
df = data.frame(a = c(1,2,3),
                b = c(4,5,6))
data.world::uploadDataFrame(connection = conn, 
                fileName="sample.csv",
                dataFrame = df, 
                dataset = "ownerid/datasetid")
```
*(Note that we only support upload a dataframe as a csv file)*

## Get Dataset
Retrieve a dataset metadata
```{r getDataset, eval = FALSE}
dataset <- getDataset(connection, dataset="user/dataset")
```
`getDataset` return a `data.world::DatasetSummaryResponse` object which is described here `https://api.data.world/v0/swagger.json`

## Download File
Download file from the latest dataset version
```{r downloadFile, eval = FALSE}
downloadFile(connection, 
             dataset="ownerid/datasetid" , 
             fileName = "file.csv",
             output = "tmp/file.csv")
```
*(Note `output` is a local file path relative to `getwd()`)*

## Download File as data frame
Download the input tabular file into an R data frame
```{r downloadFileAsDataFrame, eval = FALSE}
df <- downloadFileAsDataFrame(connection, 
                              dataset="ownerid/datasetid", 
                              fileName = "file.csv")
```
