

```{r, echo = FALSE}
library(data.world)
```

# data.world V1

## Intro 
a quick guide on how to use the r-sdk to interact with data.world platform. 
## Installation 
1. Install the latest from github
```{r, echo=TRUE, eval=FALSE}
devtools::install_github("datadotworld/data.world-r", ref = '1.0.0-sprint', build_vignettes = TRUE, force = TRUE)
library(data.world)
```
2. Configuring your client
Required initial setup step includes configurations such as data.world authentication
```{r, echo=TRUE, eval=FALSE}
data.world::configure()
```
Once this is setup, you can construct a data.world client e.g 
```{r, echo=TRUE, eval=FALSE}
datatdotworld <- data.world()
```

## SDK Usage
The main SDK expose 2 functions

```{r, echo=TRUE, eval=FALSE}
methods(class='data.world')
```
1. Declare and load a dataset dependency via `data.world::loadDataset`
```{r, echo=TRUE, eval=FALSE}
dataset <- data.world::loadDataset(data.world(), 'https://data.world/jonloyens/an-intro-to-dataworld-dataset')
```
```{r, echo=TRUE, eval=FALSE}
summary(dataset)
```
One can access tables available in a dataset via `LocalDataset$tables` properties
```{r, echo=TRUE, eval=FALSE}
names(dataset$tables)
```
```{r, echo=TRUE, eval=FALSE}
dataset$tables$datadotworldbballstats
```
```{r, echo=TRUE, eval=FALSE}
dataset$tables$datadotworldbballteam
```

2. Execute query e.g 
```{r, echo=TRUE, eval=FALSE}
queryResult <- data.world::query(data.world(), dataset = 'https://data.world/jonloyens/an-intro-to-dataworld-dataset', query = 'select * from DataDotWorldBBallTeam')
queryResult
```

## Demo
Re-running this tutorial as an R demo
```{r, echo=TRUE, eval=FALSE}
demo(package = 'data.world', topic = 'v1')
```

